Impact of seasonal forcing on a potential SARS-CoV-2 pandemic

### 2020-02-13 manuscript/cover letter.tex
manuscript update

Please receive for your consideration our manuscript “Potential impact of seasonal forcing on a SARS-CoV-2 pandemic”.

The novel virus SARS-CoV-2 has spread at an alarming pace in Wuhan and the Hubei province, and with exported cases and small transitional cluster reported globally. The WHO declared Jan 30 the SARS-CoV-2 to be a ´´\textit{public health emergency of international concern}” and Feb 11 a ´´\textit{very grave threat for the rest of the world}”. There is much uncertainty on how the epidemic will unfold and what preventive measure to undertake. In this manuscript, we show that a decrease in case number incidence during the coming months might not only be an effect of preventive measures, but also seasonal variation of the transmissibility.

Drawing on ten years data on the circulation of the “common” coronaviruses (229E, OC43, NL63, HKU1) in Stockholm, Sweden, we model the effect of possible seasonal forcing on SARS-CoV-2. Since many parameter settings are unknown, we perform simulations for a range of plausible scenarios. In these simulations, we find that in the northern hemisphere a peak might occur either in the first half of 2020, in winter 2020/2021, or that there might be two peaks with substantial regional variation.
\Jan{Robert and I discussed. Isn't already too late for a peak in the first half of 2020? Can we expect a first wave late in 2020 and a second in 2021?}

@@

We hope that you will consider our manuscript for publication through the rapid communication track and look forward to
your correspondence.

>>>>

Please receive for your consideration our manuscript “Potential impact of seasonal forcing on a SARS-CoV-2 pandemic.”

The novel virus SARS-CoV-2 has spread at an alarming pace in Wuhan and Hubei province, with exported cases and small transmission clusters reported globally. The WHO declared SARS-CoV-2 to be a ``\textit{public health emergency of international concern}" (Jan 30) and ``\textit{very grave threat for the rest of the world}” (Feb 11). There is much uncertainty on how the epidemic will unfold and what preventive measure to undertake. In this manuscript, we show that a decrease in case number incidence during the coming months might not only be an effect of preventive measures, but also seasonal variation of the transmissibility.

Drawing on ten years data on the circulation of the “common” coronaviruses (229E, OC43, NL63, HKU1) in Stockholm, Sweden, we model the effect of possible seasonal forcing on SARS-CoV-2. Since many parameter settings are unknown, we perform simulations for a range of plausible scenarios. In these simulations, we find that in the northern hemisphere a peak might occur either in the first half of 2020, in winter 2020/2021, or as two peaks with substantial regional variation.

@@

We hope that you will consider our manuscript for publication through the rapid communication track and look forward to
your correspondence. We use the new name SARS-CoV-2 as suggested by ICTV, but would be happy to change to COVID-19 if needed.
